---
title: "R STATISTICAL TESTS"
output: html_notebook
---

# 1. t-test

**one-sample t-test**


```{r}

# Generate sample data
apple_weights <- c(145, 152, 147, 155, 148)

# Perform one-sample t`-test
t.test(apple_weights, mu = 150)

```

**two-sample t-test**

```{r}
# Generate sample data
heights_male <- c(170, 175, 180, 178, 172)
heights_female <- c(160, 165, 168, 162, 170)

# Perform two-sample t-test
t.test(heights_male, heights_female)


```

**Independent t-test**

```{r}
# Generate sample data
scores_school_A <- c(85, 90, 88, 92, 87)
scores_school_B <- c(82, 88, 85, 86, 90)

# Perform independent samples t-test
t.test(scores_school_A, scores_school_B)


```

**paired t-test(dependent t-test)**

```{r}

# Generate sample data
before_medication <- c(120, 125, 118, 122, 123)
after_medication <- c(115, 120, 118, 119, 121)

# Perform paired samples t-test
t.test(before_medication, after_medication, paired = TRUE)

```

# 2. ANOVA (Analysis of Variance)

**one-way ANOVA**

```{r}
# Generate sample data
method_a_scores <- c(85, 90, 88, 92, 87)
method_b_scores <- c(82, 88, 85, 86, 90)
method_c_scores <- c(78, 84, 80, 82, 85)

# Perform one-way ANOVA
anova_result <- aov(c(method_a_scores, method_b_scores, method_c_scores) ~ c(rep("Method A", 5), rep("Method B", 5), rep("Method C", 5)))
summary(anova_result)


```

**two-way ANOVA**

```{r}

# Generate sample data
set.seed(123) # for reproducibility
gender <- factor(rep(c("Male", "Female"), each = 50))
treatment <- factor(rep(c("Drug A", "Drug B"), each = 50))
blood_pressure <- c(rnorm(50, mean = 120, sd = 10),
                    rnorm(50, mean = 130, sd = 15),
                    rnorm(50, mean = 125, sd = 12),
                    rnorm(50, mean = 135, sd = 14))

# Create a data frame
data <- data.frame(gender, treatment, blood_pressure)

# Perform two-way ANOVA
anova_result <- aov(blood_pressure ~ gender + treatment + gender:treatment, data = data)

# Summarize the ANOVA results
summary(anova_result)



```

**repeated measures ANOVA**

```{r}
# Generate sample data
before <- c(25, 28, 30, 22, 27)
immediately_after <- c(20, 18, 24, 19, 21)
one_hour_after <- c(22, 19, 18, 20, 23)

# Perform repeated measures ANOVA
anova_result <- aov(c(before, immediately_after, one_hour_after) ~ c(rep("Before", 5), rep("Immediately After", 5), rep("One Hour After", 5)))

# Get ANOVA table
anova_table <- summary(anova_result)[[1]]
# Extract relevant information
p_value <- anova_table[["Pr(>F)"]]
# Create a table manually
result_table <- data.frame(
  Source = rownames(anova_table),
  Df = anova_table[,"Df"],
  Sum_Sq = anova_table[,"Sum Sq"],
  Mean_Sq = anova_table[,"Mean Sq"],
  F_value = anova_table[,"F value"],
  P_value = p_value
)
# display the table
print(result_table)
```

# 3. Chi-squared test

**Chi-squared test of independence**

```{r}
# Create contingency table
gender <- c("Male", "Male", "Female", "Female", "Male", "Female", "Male", "Female", "Female", "Male")
voting_preference <- c("Democrat", "Republican", "Democrat", "Independent", "Republican", "Republican", "Democrat", "Democrat", "Independent", "Republican")
table_data <- table(gender, voting_preference)

# Perform Chi-squared test of independence
chi_squared_test <- chisq.test(table_data)
chi_squared_test

```

**Chi-squared goodness-of-fit**

```{r}
# Create observed and expected frequency vectors
observed_eye_colors <- c(120, 180, 90, 30)
expected_eye_colors <- c(150, 170, 80, 20)

# Perform Chi-squared goodness-of-fit test
chi_squared_test <- chisq.test(observed_eye_colors, p = expected_eye_colors / sum(expected_eye_colors))
chi_squared_test


```

# 4. Correlation test

**Pearson correlation test**

```{r}
# Generate sample data
hours_studied <- c(4, 6, 3, 5, 7, 6, 8, 5, 3, 9)
exam_scores <- c(65, 70, 60, 75, 80, 70, 85, 70, 55, 90)

# Perform Pearson correlation test
pearson_correlation <- cor.test(hours_studied, exam_scores, method = "pearson")
pearson_correlation


```

**Spearman correlation test**

```{r}
# Generate sample data
exam1_ranks <- c(1, 2, 3, 4, 5)
exam2_ranks <- c(2, 3, 5, 1, 4)

# Perform Spearman correlation test
spearman_correlation <- cor.test(exam1_ranks, exam2_ranks, method = "spearman")
spearman_correlation


```

**Kendall correlation test**

```{r}
# Generate sample data
exam1_ranks <- c(1, 2, 3, 4, 5)
exam2_ranks <- c(2, 3, 5, 1, 4)

# Perform Kendall correlation test
kendall_correlation <- cor.test(exam1_ranks, exam2_ranks, method = "kendall")
kendall_correlation


```

# 5. Linear regression

**Simple Linear Regression**

```{r}
# Sample data for simple linear regression
hours_studied <- c(4, 6, 3, 5, 7, 6, 8, 5, 3, 9)
exam_scores <- c(65, 70, 60, 75, 80, 70, 85, 70, 55, 90)

# Fit simple linear regression model
simple_linear_model <- lm(exam_scores ~ hours_studied)

# Summary of the simple linear regression model
summary(simple_linear_model)


```

**Multiple Linear Regression**

```{r}
# Generate sample data
area <- c(1000, 1500, 1200, 1800, 2000)
bedrooms <- c(2, 3, 2, 4, 3)
location <- c(1, 2, 2, 3, 1)
price <- c(103.2, 209.23, 289.23, 378.34, 100.23)

# Fit multiple linear regression model
multiple_linear_model <- lm(price ~ area + bedrooms + location)

# Summary of the multiple linear regression model
summary(multiple_linear_model)


```

**Polynomial Regression**

```{r}

# Sample data for polynomial regression
advertising <- c(100, 200, 300, 400, 500)
sales <- c(50, 80, 120, 150, 200)

# Fit polynomial regression model
polynomial_model <- lm(sales ~ poly(advertising, 2))

# Summary of the polynomial regression model
summary(polynomial_model)


```

# 6. Logistic regression

**Binary Logistic Regression**

```{r}

# Sample data for binary logistic regression
# Assuming 1 represents spam and 0 represents not spam
spam <- c(1, 0, 1, 0, 1)
subject_length <- c(10, 20, 15, 25, 12)

# Fit binary logistic regression model
binary_logistic_model <- glm(spam ~ subject_length, family = binomial(link = "logit"))

# Summary of the binary logistic regression model
summary(binary_logistic_model)

```

**Multinomial Logistic Regression**

```{r}

# Load the required package
library(nnet)

# Sample data for multinomial logistic regression
crop_type <- c("wheat", "rice", "maize", "wheat", "rice")
ph_level <- c(6.5, 7.2, 5.8, 6.9, 6.0)

# Convert crop_type to a factor
crop_type <- factor(crop_type)

# Create a data frame
data <- data.frame(crop_type, ph_level)

# Fit multinomial logistic regression model
multinomial_logistic_model <- multinom(crop_type ~ ph_level, data = data)

# Summary of the multinomial logistic regression model
summary(multinomial_logistic_model)



```

**Ordinal Logistic Regression**

```{r}
# Load the required package
library(MASS)

# Sample data for ordinal logistic regression
satisfaction <- ordered(c("low", "medium", "high", "low", "high"))
service_quality <- c(3, 4, 2, 5, 3)

# Fit ordinal logistic regression model
ordinal_logistic_model <- polr(satisfaction ~ service_quality, method = "logistic")

# Summary of the ordinal logistic regression model
summary(ordinal_logistic_model)

```
